### 2022-02-06 ms_body/manuscript.tex
Add BibTex references to MS

Sutton, G. F.\textsuperscript{1}\(\dag\); Martin, G.D.
\textsuperscript{1,2}

@@

Ecological models are important tools to support and guide the
development and implementation of environmental policies and management
programs (\textbf{Addison2013?}; \textbf{Schuwirth2019?}). For example,
ecological models can be used for planning conservation programs
(\textbf{Guisan2013?}), predicting the establishment of invasive species
(\textbf{Martin2020?}), implementing biological control programmes
(\textbf{Mukherjee2021?}), or forecasting species responses to
environmental change (\textbf{Bocedi2014?}). Species distribution models
(SDM's) are an example of ecological models that have become
increasingly popular in recent years (\textbf{Elith2009?}). SDM's
typically take the form of correlative or mechanistic models that relate
species presence/absences to environmental covariates to map habitats
that may be suitable for a taxon (\textbf{Elith2011?}). The Maximum
Entropy Species Distribution Model (hereafter `MaxEnt')
(\textbf{Phillips2017?}) is among the most popular methods for species
distribution modeling and has been shown to perform well compared to
alternative modeling algorithms (\textbf{Elith2006?}). It uses maximum
entropy to distinguish between environmental conditions where the focal
taxon is present from environmental conditions at background locations
where the taxons' presence is absent (or assumed absent)
(\textbf{Elith2011?}).

Despite its popularity, the usefulness of MaxEnt has been questioned
when used to generate predictions for novel geographic regions
(transferred in space) or time periods (transferred in time) from those
for which the data used to build the model were collected
(\textbf{Yates2018?}; \textbf{Blasi2021?}). To date, studies assessing
transferability have produced contrasting results (\textbf{Randin2006?};
\textbf{Bahn2013?}; \textbf{Capinha2018?}). Model transferability has
typically been assessed by cross validation, whereby a fraction of the
data is withheld to calibrate the model (i.e.~training data) and the
remaining data is retained to validate the model (i.e.~the testing data)
(\textbf{Wenger2012?}). However, cross validation may not provide an
estimate of model transferability as a selected test sample may not
provide an unbiased estimate of the full data set, and therefore may not
reflect the heterogeneity that may exist in the underlying data in space
or time (see \textbf{Wenger2012?}). Ideally, a direct test of model
transferability should use an independent dataset of species
presence/absences and environmental covariates (\textbf{Fielding1997?};
\textbf{West2016?}). Unfortunately, very few studies, to date, have
tested model transferability using independent datasets (e.g.
\textbf{Costa2010?}; \textbf{Rebelo2010?}), which may have limited our
understanding of how useful SDM's can be when transferred in space or
time. Issues with model transferability may lead to misleading
conclusions, which may hinder scientists abilities to address important
ecological questions (e.g.~where could an invasive species establish?,
where should a conservation priority be translocated?)
(\textbf{Merow2013?}).

Model complexity can affect uncertainty of model transferability in
space and time (\textbf{Warren2014?}). Model complexity may be affected
by numerous parameters, notably the feature classes and regularisation
(beta) multiplier settings used to specify the MaxEnt model
(\textbf{Elith2011?}; \textbf{Merow2013?}). MaxEnt allows users to
specify combinations of different feature classes that modify the shape
and complexity of species-environment functions. Moreover, MaxEnt uses a
regularisation multiplier (a form of L\textsubscript{1}-regularisation)
to balance model complexity with the model fit, penalising unnecessarily
complex models (\textbf{Merow2013?}). The default MaxEnt settings use
all feature classes in model fitting except `threshold' features
(depending on sample size of training data) and a regularisation
multiplier of 1 (\textbf{Phillips2017?}). Several recent studies have
highlighted the importance of species-specific tuning of feature classes
and regularisation multipliers for optimising model complexity
(\textbf{Shcheglovitova2013?}; \textbf{Radosavljevic2014?}), with
default settings often producing models that are overly complex, and
thus, may perform poorly when transferred (\textbf{Elith2011?};
\textbf{Merow2013?}). Despite the potential implications, surprisingly
few studies use user-defined feature class and regularisation settings
(e.g. \textbf{Sutton2019b?}), and as such, little is known about how
model settings configurations may influence model transferability
(\textbf{Moreno-Amat2015?}; \textbf{Low2021?}).

In this study, we performed a retrospective assessment of the
transferability of SDM's built using MaxEnt (\textbf{Phillips2017?}).
Here, we modelled the potential climatic suitability of an insect that
is native to Australia, and that has been introduced into South Africa
as a biological control agent. The insect, \emph{Dasineura rubiformis}
Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from
Australia in 2013, to curb the spread of the invasive Australian
\emph{Acacia mearnsii} De Wild (\textbf{Impson2013?}). Extensive field
surveys have been performed in Australia (characterizing the native
distribution of \emph{D. rubiformis}) (\textbf{Adair2004?}) and
long-term post-release monitoring has been performed in South Africa to
measure establishment rates of the insect, following experimental
releases of the insect (Impson et al., 2021). As such, the availability
of independent testing data from a separate continent to the data used
to generate the models makes this an ideal study system for assessing
the transferability of SDM's.

@@

A total of 103 native-range records for \emph{D. rubiformis} were
obtained from (\textbf{Adair2004?}). A total of 452 invaded-range
records to evaluate model transferability were taken from Impson,
Kleinjan and Hoffmann (unpublished). Distribution maps for \emph{D.
rubiformis} for both native and invaded range are provided in Fig. 1.

Spatial autocorrelation is an important factor that may affect model
outputs. Filtering of species occurrence data may limit the inherent
biases in the data and improve model quality (\textbf{Veloz2009?}). To
avoid pseudo-replication, only one occurrence record per 2.5 minute grid
cell was used for model calibration. Species occurrence datasets were
thinned using the `spThin' package (\textbf{Aiello-Lammens2015?}), and
spatial autocorrelation analyses were performed using the `ecospat'
package (\textbf{DiCola2017?}).

@@

Climate data were obtained by downloading the standard set of 19
bioclimatic variables from the WorldClim ver. 1.4 database
(\textbf{Hijmans2005?}) (data available at:
www.worldclim.org/download.html). This dataset is representative of
annual and seasonal means and variation of temperature and precipitation
metrics averaged over the 1950--2000 time period (current climate) at a
5 arc minute resolution.

To reduce multicollinearity between environmental predictors and thereby
decrease the risk of calibrating overfit models, Pearson's correlation
coefficients were computed for all pairs of predictors, whereby
predictors which were highly correlated (\textbar{}\emph{r}\textbar{}
\textgreater{} 0.85) were excluded from the final predictor set
(\textbf{Capinha2011?}). The reduced set of environmental predictors
consisted of seven climatic variables, including: bio1 -- annual mean
temperature, bio2 -- mean diurnal temperature range (mean of monthly
{[}max temp -- min temp{]}, bio3 -- isothermality, bio9 -- mean
temperature of the driest quarter, bio12 -- mean annual precipitation,
bio14 -- precipitation of the driest month and bio16 -- precipitation of
the wettest quarter (see (\textbf{Hijmans2005?}) for further details).

@@

Maxent was implemented in the \texttt{dismo} package in \texttt{R}
(\textbf{hijmans2017package?}). MaxEnt was selected as it consistently
outperforms other modelling algorithms (\textbf{Wisz2008?}). Given that
MaxEnt is a presence-only modelling algorithm, model calibration
requires a user-defined geographic background to sample the climate of
representative grid cells where the focal species is absent
(i.e.~background points). Background definition can have a significant
effect on model output (\textbf{VanDerWal2009?}). The background should
ideally represent the geographic areas available to the focal species,
omitting areas where species absence is due to historical factors,
dispersal constraints and/or biotic interactions (\textbf{Sanin2018?}).
Following (\textbf{Webber2011?}), we defined the model background using
the Koppen-Geiger climate classification (Available at:
\url{http://koeppen-geiger.vu-wien.ac.at}). Only Koppen-Geiger climate
zones that contained at least one native-range occurrence record for
\emph{D. rubiformis} were used as the background area from which
pseudo-absences were drawn for model calibration. Koppen-Geiger climate
zones were intersected using the \texttt{extract} function from the
\texttt{R} package \texttt{raster} (\textbf{hijmans2017package?}).

@@

Model performance and optimal settings configurations were assessed
using multiple metrics that reflected aspects of model (1)
discriminatory ability (AUC\textsubscript{test}), (2) overfitting
(AUC\textsubscript{diff}), (3) omission rates (OR\textsubscript{10}),
and (4) overall parsimony (AICc) (\textbf{Low2021?}).

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  AUC\textsubscript{test} assesses the models ability to discriminate
  between predicted presence/absence at withheld portions of the data
  used to test the model versus background points. The area under the
  receiver operating characteristic curve (AUC) is one of the most
  popular metrics used to evaluate MaxEnt models. An AUC of less than
  0.8 is considered a poor model, between 0.8 and 0.9 is a fair model,
  between 0.9 and 0.995 a good model, and \textgreater{} 0.995 an
  excellent model (\textbf{Fielding1997?}). Thus, higher
  AUC\textsubscript{test} values indicate increased ability to
  discriminate between training and background points.
\item
  AUC\textsubscript{diff} calculates the difference between AUC values
  calculated on training points only (AUC\textsubscript{train}) and
  AUC\textsubscript{test} using cross-validation (\textbf{Warren2011?}).
  Thus, higher AUC\textsubscript{diff} values indicate whether the
  MaxEnt model is overfit on the training data, and thus, performs
  poorly when evaluated against testing points.
\item
  OR\textsubscript{10} is the proportion of testing points that are not
  predicted to fall within the projected model surface once the model is
  converted into a binary prediction output (\textbf{Boria2014?}).
  Overfit models have omission rates higher than the theoretical
  expectation for the threshold applied (\textbf{Shcheglovitova2013?}).
  OR10 sets the binary prediction threshold at a value that excludes the
  10\% of the calibration localities from the model with the lowest
  prediction values, and therefore has an expected omission rate of 0.10
  (\textbf{Boria2014?}). As such, the OR\textsubscript{10} criterion
  selected models calibrated with MaxEnt settings which best
  approximated the expected 0.10 omission rate. Models with omission
  rates increasingly higher than the expected value were considered as
  more overfit (\textbf{Boria2017?}).
\item
  Lastly, optimal model settings were determined by selecting model
  configurations which produced the lowest value for the Akaike
  Information Criterion corrected for small sample sizes (AICc)
  (i.e.~AICc=0; following (\textbf{Muscarella2014?})). The AICc
  criterion simultaneously scores models according to their complexity
  and goodness-of-fit, whereby models with the lowest AICc are selected
  as the best models. AICc was used as the primary evaluation metric as
  it is calculated using MaxEnt models built using the entire species
  occurrence dataset, unlike AUC and OR\textsubscript{10} (and numerous
  other metrics frequently used for model evaluation) which may be
  spatially biased due to the partitioning of the species occurrence
  dataset into training and evaluation sets (\textbf{Sanin2018?}).
\end{enumerate}

@@

Several authors have raised concerns over the use of correlative SDM's,
such as MaxEnt, when projected into new geographic regions or time
periods due to issues with extrapolation into novel/non-analogous
climate (\textbf{Elith2009?}; \textbf{Elith2011?}; \textbf{Yates2018?}).
To address these concerns, Multivariate environmental similarity
surfaces (hereafter `MESS') (\textbf{Elith2010?}) were computed to
assess whether MaxEnt models were extrapolating or interpolating. MESS
analyses measure the similarity of any given point to a set of reference
points (reference points were defined as only the species presence
points used to calibrate models following \textbf{Kriticos2014?}).
Negative MESS values indicate geographic regions outside the range of
climate variables used to calibrate the model (i.e.~extrapolation space
or MESS-), while MESS values between 0 and 100 indicate geographic
regions inside the range of climatic variables used to calibrate the
model (i.e.~interpolation space or MESS+). MESS maps are used as a
measure of prediction uncertainty or caution against inferences in
extrapolation space (\textbf{Elith2010?}).

@@

To evaluate the predictive performance and transferability of these
models, we evaluated whether MaxEnt climatic suitability scores were
correlated with increased probabilities of \emph{D. rubiformis}
establishment in South Africa. To do so, MaxEnt scores were specified as
a continuous fixed effect and whether \emph{D. rubiformis} was recorded
or not (present/absent) as a boolean response variable in a logistic
general linear model (GLM), using a logit link function. MaxEnt scores,
bounded between 0 (not climatically suitable) and 1 (highly climatically
suitable) indicating climatic similarity between the native range
occupied by \emph{D. rubiformis} in Australia and sites where it was
released as a biological control agent in South Africa, were derived by
projecting MaxEnt suitability rasters from models calibrated using
native-range GPS records only (Australian native-range records) over
South Africa, and extracting the MaxEnt suitability scores for each site
where \emph{D. rubiformis} was surveyed in South Africa using the
\texttt{extract} function from the \texttt{R} package \texttt{raster}
(\textbf{hijmans2017package?}). GLM's were specified using the
\texttt{rms} package in \texttt{R} (\textbf{Harrell2017package?}).
Likelihood Ratio Tests (LRT) were performed to determine whether there
was a statistical correlation between MaxEnt scores and the probability
of \emph{D. rubiformis} establishment. A statistically significant,
positive correlation between MaxEnt scores and insect establishment
rates would provide support that the SDM's were useful and predictive of
insect establishment, even when the insect was transferred in space.
Out-of-sample AUC was calculated using the \texttt{rms} package, using
500 bootstrap replicates (\textbf{Harrell2017package?}) to quantify the
predictive performance of MaxEnt models, whereby AUC values less than
0.80 are typically considered poor fitting models (lacks predictive
power) (\textbf{Fielding1997?}).

A similar approach was adopted to test if each occurrence record in
South Africa was classified as being in interpolation (within the
climate space occupied by the insect in its native range) versus
extrapolation (outside the climate space occupied by the insect in its
native range) was predictive of \emph{D. rubiformis} establishment.
Interpolation/extrapolation were quantified using MESS maps using the
\texttt{mess} function from the \texttt{dismo} package in \texttt{R}
(\textbf{hijmans2017package?}). Interpolation versus extrapolation was
fit as a categorical predictor and whether \emph{D. rubiformis} was
recorded or not (present/absent) as a boolean response variable in a
logistic GLM, using a logit link function.

All modelling and statistical analyses were conducted in \texttt{R} ver.
4.0.3 (\textbf{RCoreTeam2020?}).

@@

Species distribution models (SDM's) are becoming increasingly important
tools to support and guide the development and implementation of
environmental policies and management programs (\textbf{Addison2013?};
\textbf{Schuwirth2019?}). The usefulness of SDM's is dependent on model
performance, which is typically measured by its ability to correctly
predict where a species is established and its capacity to distinguish
between sites where the species is established versus a site where the
species is absent (or assumed to be absent) (\textbf{Smith2021?}).
However, a number of authors have called in question the usefulness of
SDM's, notably when models are transferred in space or time
(\textbf{Yates2018?}; \textbf{Blasi2021?}). Poor transferability may
lead to erroneous model predictions that may have serious negative
consequences for the application of the model, and resulting policies
and management programmes (Soultan and Safi, 2017). Here, we report on
the validation of a SDM for an insect, (\emph{D. rubiformis}), native to
Australia, that has been intentionally introduced as a biological
control agent into South Africa to reduce the invasiveness of its
host-plant, \emph{A. mearnsii} (Impson et al., 2013). The the
availability of independent testing data from South Africa, a separate
continent to the data used to generate the models (Australia), makes
this an ideal study system for assessing the transferability of SDM's.
This study represents one of only a handful of studies to validate SDM's
using independent testing data (\textbf{Costa2010?};
\textbf{Rebelo2010?}; \textbf{West2016?}; \textbf{Smith2021?}).

Our results indicated that MaxEnt SDM's were able to provide robust
estimates of climatic suitability for \emph{D. rubiformis} in South
Africa. AUC values greater than 0.80 are typically considered good
fitting models (having high predictive power) (\textbf{Fielding1997?}).
When MaxEnt models, calibrated using native-range records for \emph{D.
rubiformis} from Australia, were projected over South Africa, they had
AUC scores of 0.83 - 0.84, indicative of high power to discriminate
between sites where \emph{D. rubiformis} had established in South Africa
or not. Additionally, we showed that the probability that \emph{D.
rubiformis} was recorded at a field site in South Africa increased with
higher MaxEnt climatic suitability scores. This result demonstrates the
usefulness of the MaxEnt SDM and that, at least for our case-study,
SDM's can be transferred into novel geographic regions, and provide
valuable information to guide policy making and the development of
management programmes. For example, biological control programmes such
as the programme in which \emph{D. rubiformis} is used to curtail the
spread of the highly invasive weed \emph{A. mearnsii} expend many
resources (time and money) to mass-rear insect populations that are then
released into the field. The positive correlation between MaxEnt
suitability scores and \emph{D. rubiformis} establishment means that
researchers can develop and implement mass-rearing and experimental
release programmes for the insect that prioritise releasing the insect
at sites that are highly climatically suitable, and thus, more likely
the insect will establish viable populations. This may assist in
improving the level of control over the target weed and increasing the
efficiency of resources expended by increasing the chances of
establishing viable insect populations.

Our study adds to a growing body of literature indicating that SDM model
complexity affects model performance and transferability when being
projected into novel geographic regions. While the default MaxEnt model
had the highest predictive power in the training area (Australia), this
model performed extremely poorly when transferred over South Africa. In
contrast, the four MaxEnt models calibrated with tuned settings showed
high predictive power when transferred. The poor transferability of the
Default model is likely explained by default settings overfitting the
model on the training points (\textbf{Warren2014?}). This finding
corroborates other studies showing that high predictive power in the
model training area is not a guarantee of transferability
(\textbf{Warren2014?}; \textbf{Duque-Lazo2016?}). We echo that
recommendations of (\textbf{Shcheglovitova2013?}) and
(\textbf{Muscarella2014?}) that model tuning should be performed to
estimate optimal model complexity when transferring MaxEnt models in
space and/or time in order to maximise their usefulness.

>>>>

Sutton, G. F.\textsuperscript{1}\(\dag\); Martin,
G.D.\textsuperscript{1,2}

@@

Ecological models are important tools to support and guide the
development and implementation of environmental policies and management
programs (Addison et al., 2013; Schuwirth et al., 2019). For example,
ecological models can be used for planning conservation programs (Guisan
et al., 2013), predicting the establishment of invasive species (Martin
et al., 2020), implementing biological control programmes (Mukherjee et
al., 2021), or forecasting species responses to environmental change
(Bocedi et al., 2014). Species distribution models (SDM's) are an
example of ecological models that have become increasingly popular in
recent years (Elith and Leathwick, 2009). SDM's typically take the form
of correlative or mechanistic models that relate species
presence/absences to environmental covariates to map habitats that may
be suitable for a taxon (Elith et al., 2011). The Maximum Entropy
Species Distribution Model (hereafter `MaxEnt') (Phillips et al., 2017)
is among the most popular methods for species distribution modeling and
has been shown to perform well compared to alternative modeling
algorithms (Elith et al., 2006). It uses maximum entropy to distinguish
between environmental conditions where the focal taxon is present from
environmental conditions at background locations where the taxons'
presence is absent (or assumed absent) (Elith et al., 2011).

Despite its popularity, the usefulness of MaxEnt has been questioned
when used to generate predictions for novel geographic regions
(transferred in space) or time periods (transferred in time) from those
for which the data used to build the model were collected (Blasi et al.,
2021; Yates et al., 2018). To date, studies assessing transferability
have produced contrasting results (Bahn and McGill, 2013; Capinha et
al., 2018; Randin et al., 2006). Model transferability has typically
been assessed by cross validation, whereby a fraction of the data is
withheld to calibrate the model (i.e.~training data) and the remaining
data is retained to validate the model (i.e.~the testing data) (Wenger
and Olden, 2012). However, cross validation may not provide an estimate
of model transferability as a selected test sample may not provide an
unbiased estimate of the full data set, and therefore may not reflect
the heterogeneity that may exist in the underlying data in space or time
(see Wenger and Olden, 2012). Ideally, a direct test of model
transferability should use an independent dataset of species
presence/absences and environmental covariates (Fielding and Bell, 1997;
West et al., 2016). Unfortunately, very few studies, to date, have
tested model transferability using independent datasets (e.g. Costa et
al., 2010; Rebelo and Jones, 2010), which may have limited our
understanding of how useful SDM's can be when transferred in space or
time. Issues with model transferability may lead to misleading
conclusions, which may hinder scientists abilities to address important
ecological questions (e.g.~where could an invasive species establish?,
where should a conservation priority be translocated?) (Merow et al.,
2013).

Model complexity can affect uncertainty of model transferability in
space and time (Warren et al., 2014). Model complexity may be affected
by numerous parameters, notably the feature classes and regularisation
(beta) multiplier settings used to specify the MaxEnt model (Elith et
al., 2011; Merow et al., 2013). MaxEnt allows users to specify
combinations of different feature classes that modify the shape and
complexity of species-environment functions. Moreover, MaxEnt uses a
regularisation multiplier (a form of L\textsubscript{1}-regularisation)
to balance model complexity with the model fit, penalising unnecessarily
complex models (Merow et al., 2013). The default MaxEnt settings use all
feature classes in model fitting except `threshold' features (depending
on sample size of training data) and a regularisation multiplier of 1
(Phillips et al., 2017). Several recent studies have highlighted the
importance of species-specific tuning of feature classes and
regularisation multipliers for optimising model complexity
(Radosavljevic and Anderson, 2014; Shcheglovitova and Anderson, 2013),
with default settings often producing models that are overly complex,
and thus, may perform poorly when transferred (Elith et al., 2011; Merow
et al., 2013). Despite the potential implications, surprisingly few
studies use user-defined feature class and regularisation settings (e.g.
Sutton, 2019), and as such, little is known about how model settings
configurations may influence model transferability (Low et al., 2021;
Moreno-Amat et al., 2015).

In this study, we performed a retrospective assessment of the
transferability of SDM's built using MaxEnt (Phillips et al., 2017).
Here, we modelled the potential climatic suitability of an insect that
is native to Australia, and that has been introduced into South Africa
as a biological control agent. The insect, \emph{Dasineura rubiformis}
Kolesik (Diptera: Cecidomyiidae) was introduced into South Africa from
Australia in 2013, to curb the spread of the invasive Australian
\emph{Acacia mearnsii} De Wild (Impson et al., 2013). Extensive field
surveys have been performed in Australia (characterizing the native
distribution of \emph{D. rubiformis}) (Adair, 2004) and long-term
post-release monitoring has been performed in South Africa to measure
establishment rates of the insect, following experimental releases of
the insect (Impson et al., 2021). As such, the availability of
independent testing data from a separate continent to the data used to
generate the models makes this an ideal study system for assessing the
transferability of SDM's.

@@

A total of 103 native-range records for \emph{D. rubiformis} were
obtained from (Adair, 2004). A total of 452 invaded-range records to
evaluate model transferability were taken from Impson, Kleinjan and
Hoffmann (unpublished). Distribution maps for \emph{D. rubiformis} for
both native and invaded range are provided in Fig. 1.

Spatial autocorrelation is an important factor that may affect model
outputs. Filtering of species occurrence data may limit the inherent
biases in the data and improve model quality (Veloz, 2009). To avoid
pseudo-replication, only one occurrence record per 2.5 minute grid cell
was used for model calibration. Species occurrence datasets were thinned
using the `spThin' package (Aiello-Lammens et al., 2015), and spatial
autocorrelation analyses were performed using the `ecospat' package (Di
Cola et al., 2017).

@@

Climate data were obtained by downloading the standard set of 19
bioclimatic variables from the WorldClim ver. 1.4 database (Hijmans et
al., 2005) (data available at: www.worldclim.org/download.html). This
dataset is representative of annual and seasonal means and variation of
temperature and precipitation metrics averaged over the 1950--2000 time
period (current climate) at a 5 arc minute resolution.

To reduce multicollinearity between environmental predictors and thereby
decrease the risk of calibrating overfit models, Pearson's correlation
coefficients were computed for all pairs of predictors, whereby
predictors which were highly correlated (\textbar{}\emph{r}\textbar{}
\textgreater{} 0.85) were excluded from the final predictor set (Capinha
and Anastácio, 2011). The reduced set of environmental predictors
consisted of seven climatic variables, including: bio1 -- annual mean
temperature, bio2 -- mean diurnal temperature range (mean of monthly
{[}max temp -- min temp{]}, bio3 -- isothermality, bio9 -- mean
temperature of the driest quarter, bio12 -- mean annual precipitation,
bio14 -- precipitation of the driest month and bio16 -- precipitation of
the wettest quarter (see Hijmans et al. (2005) for further details).

@@

Maxent was implemented in the \texttt{dismo} package in \texttt{R}
(Hijmans et al., 2017). MaxEnt was selected as it consistently
outperforms other modelling algorithms (Wisz et al., 2008). Given that
MaxEnt is a presence-only modelling algorithm, model calibration
requires a user-defined geographic background to sample the climate of
representative grid cells where the focal species is absent
(i.e.~background points). Background definition can have a significant
effect on model output (VanDerWal et al., 2009). The background should
ideally represent the geographic areas available to the focal species,
omitting areas where species absence is due to historical factors,
dispersal constraints and/or biotic interactions (Sanín and Anderson,
2018). Following Webber et al. (2011), we defined the model background
using the Koppen-Geiger climate classification (Available at:
\url{http://koeppen-geiger.vu-wien.ac.at}). Only Koppen-Geiger climate
zones that contained at least one native-range occurrence record for
\emph{D. rubiformis} were used as the background area from which
pseudo-absences were drawn for model calibration. Koppen-Geiger climate
zones were intersected using the \texttt{extract} function from the
\texttt{R} package \texttt{raster} (Hijmans et al., 2017).

@@

Model performance and optimal settings configurations were assessed
using multiple metrics that reflected aspects of model (1)
discriminatory ability (AUC\textsubscript{test}), (2) overfitting
(AUC\textsubscript{diff}), (3) omission rates (OR\textsubscript{10}),
and (4) overall parsimony (AICc) (Low et al., 2021).

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  AUC\textsubscript{test} assesses the models ability to discriminate
  between predicted presence/absence at withheld portions of the data
  used to test the model versus background points. The area under the
  receiver operating characteristic curve (AUC) is one of the most
  popular metrics used to evaluate MaxEnt models. An AUC of less than
  0.8 is considered a poor model, between 0.8 and 0.9 is a fair model,
  between 0.9 and 0.995 a good model, and \textgreater{} 0.995 an
  excellent model (Fielding and Bell, 1997). Thus, higher
  AUC\textsubscript{test} values indicate increased ability to
  discriminate between training and background points.
\item
  AUC\textsubscript{diff} calculates the difference between AUC values
  calculated on training points only (AUC\textsubscript{train}) and
  AUC\textsubscript{test} using cross-validation (Warren and Seifert,
  2011). Thus, higher AUC\textsubscript{diff} values indicate whether
  the MaxEnt model is overfit on the training data, and thus, performs
  poorly when evaluated against testing points.
\item
  OR\textsubscript{10} is the proportion of testing points that are not
  predicted to fall within the projected model surface once the model is
  converted into a binary prediction output (Boria et al., 2014).
  Overfit models have omission rates higher than the theoretical
  expectation for the threshold applied (Shcheglovitova and Anderson,
  2013). OR10 sets the binary prediction threshold at a value that
  excludes the 10\% of the calibration localities from the model with
  the lowest prediction values, and therefore has an expected omission
  rate of 0.10 (Boria et al., 2014). As such, the OR\textsubscript{10}
  criterion selected models calibrated with MaxEnt settings which best
  approximated the expected 0.10 omission rate. Models with omission
  rates increasingly higher than the expected value were considered as
  more overfit (Boria et al., 2017).
\item
  Lastly, optimal model settings were determined by selecting model
  configurations which produced the lowest value for the Akaike
  Information Criterion corrected for small sample sizes (AICc)
  (i.e.~AICc=0; following Muscarella et al. (2014)). The AICc criterion
  simultaneously scores models according to their complexity and
  goodness-of-fit, whereby models with the lowest AICc are selected as
  the best models. AICc was used as the primary evaluation metric as it
  is calculated using MaxEnt models built using the entire species
  occurrence dataset, unlike AUC and OR\textsubscript{10} (and numerous
  other metrics frequently used for model evaluation) which may be
  spatially biased due to the partitioning of the species occurrence
  dataset into training and evaluation sets (Sanín and Anderson, 2018).
\end{enumerate}

@@

Several authors have raised concerns over the use of correlative SDM's,
such as MaxEnt, when projected into new geographic regions or time
periods due to issues with extrapolation into novel/non-analogous
climate (Elith et al., 2011; Elith and Leathwick, 2009; Yates et al.,
2018). To address these concerns, Multivariate environmental similarity
surfaces (hereafter `MESS') (Elith et al., 2010) were computed to assess
whether MaxEnt models were extrapolating or interpolating. MESS analyses
measure the similarity of any given point to a set of reference points
(reference points were defined as only the species presence points used
to calibrate models following Kriticos et al., 2014). Negative MESS
values indicate geographic regions outside the range of climate
variables used to calibrate the model (i.e.~extrapolation space or
MESS-), while MESS values between 0 and 100 indicate geographic regions
inside the range of climatic variables used to calibrate the model
(i.e.~interpolation space or MESS+). MESS maps are used as a measure of
prediction uncertainty or caution against inferences in extrapolation
space (Elith et al., 2010).

@@

To evaluate the predictive performance and transferability of these
models, we evaluated whether MaxEnt climatic suitability scores were
correlated with increased probabilities of \emph{D. rubiformis}
establishment in South Africa. To do so, MaxEnt scores were specified as
a continuous fixed effect and whether \emph{D. rubiformis} was recorded
or not (present/absent) as a boolean response variable in a logistic
general linear model (GLM), using a logit link function. MaxEnt scores,
bounded between 0 (not climatically suitable) and 1 (highly climatically
suitable) indicating climatic similarity between the native range
occupied by \emph{D. rubiformis} in Australia and sites where it was
released as a biological control agent in South Africa, were derived by
projecting MaxEnt suitability rasters from models calibrated using
native-range GPS records only (Australian native-range records) over
South Africa, and extracting the MaxEnt suitability scores for each site
where \emph{D. rubiformis} was surveyed in South Africa using the
\texttt{extract} function from the \texttt{R} package \texttt{raster}
(Hijmans et al., 2017). GLM's were specified using the \texttt{rms}
package in \texttt{R} (Harrell, 2017). Likelihood Ratio Tests (LRT) were
performed to determine whether there was a statistical correlation
between MaxEnt scores and the probability of \emph{D. rubiformis}
establishment. A statistically significant, positive correlation between
MaxEnt scores and insect establishment rates would provide support that
the SDM's were useful and predictive of insect establishment, even when
the insect was transferred in space. Out-of-sample AUC was calculated
using the \texttt{rms} package, using 500 bootstrap replicates (Harrell,
2017) to quantify the predictive performance of MaxEnt models, whereby
AUC values less than 0.80 are typically considered poor fitting models
(lacks predictive power) (Fielding and Bell, 1997).

A similar approach was adopted to test if each occurrence record in
South Africa was classified as being in interpolation (within the
climate space occupied by the insect in its native range) versus
extrapolation (outside the climate space occupied by the insect in its
native range) was predictive of \emph{D. rubiformis} establishment.
Interpolation/extrapolation were quantified using MESS maps using the
\texttt{mess} function from the \texttt{dismo} package in \texttt{R}
(Hijmans et al., 2017). Interpolation versus extrapolation was fit as a
categorical predictor and whether \emph{D. rubiformis} was recorded or
not (present/absent) as a boolean response variable in a logistic GLM,
using a logit link function.

All modelling and statistical analyses were conducted in \texttt{R} ver.
4.0.3 (Team, 2020).

@@

Species distribution models (SDM's) are becoming increasingly important
tools to support and guide the development and implementation of
environmental policies and management programs (Addison et al., 2013;
Schuwirth et al., 2019). The usefulness of SDM's is dependent on model
performance, which is typically measured by its ability to correctly
predict where a species is established and its capacity to distinguish
between sites where the species is established versus a site where the
species is absent (or assumed to be absent) (Smith et al., 2021).
However, a number of authors have called in question the usefulness of
SDM's, notably when models are transferred in space or time (Blasi et
al., 2021; Yates et al., 2018). Poor transferability may lead to
erroneous model predictions that may have serious negative consequences
for the application of the model, and resulting policies and management
programmes (Soultan and Safi, 2017). Here, we report on the validation
of a SDM for an insect, (\emph{D. rubiformis}), native to Australia,
that has been intentionally introduced as a biological control agent
into South Africa to reduce the invasiveness of its host-plant, \emph{A.
mearnsii} (Impson et al., 2013). The the availability of independent
testing data from South Africa, a separate continent to the data used to
generate the models (Australia), makes this an ideal study system for
assessing the transferability of SDM's. This study represents one of
only a handful of studies to validate SDM's using independent testing
data (Costa et al., 2010; Rebelo and Jones, 2010; Smith et al., 2021;
West et al., 2016).

Our results indicated that MaxEnt SDM's were able to provide robust
estimates of climatic suitability for \emph{D. rubiformis} in South
Africa. AUC values greater than 0.80 are typically considered good
fitting models (having high predictive power) (Fielding and Bell, 1997).
When MaxEnt models, calibrated using native-range records for \emph{D.
rubiformis} from Australia, were projected over South Africa, they had
AUC scores of 0.83 - 0.84, indicative of high power to discriminate
between sites where \emph{D. rubiformis} had established in South Africa
or not. Additionally, we showed that the probability that \emph{D.
rubiformis} was recorded at a field site in South Africa increased with
higher MaxEnt climatic suitability scores. This result demonstrates the
usefulness of the MaxEnt SDM and that, at least for our case-study,
SDM's can be transferred into novel geographic regions, and provide
valuable information to guide policy making and the development of
management programmes. For example, biological control programmes such
as the programme in which \emph{D. rubiformis} is used to curtail the
spread of the highly invasive weed \emph{A. mearnsii} expend many
resources (time and money) to mass-rear insect populations that are then
released into the field. The positive correlation between MaxEnt
suitability scores and \emph{D. rubiformis} establishment means that
researchers can develop and implement mass-rearing and experimental
release programmes for the insect that prioritise releasing the insect
at sites that are highly climatically suitable, and thus, more likely
the insect will establish viable populations. This may assist in
improving the level of control over the target weed and increasing the
efficiency of resources expended by increasing the chances of
establishing viable insect populations.

Our study adds to a growing body of literature indicating that SDM model
complexity affects model performance and transferability when being
projected into novel geographic regions. While the default MaxEnt model
had the highest predictive power in the training area (Australia), this
model performed extremely poorly when transferred over South Africa. In
contrast, the four MaxEnt models calibrated with tuned settings showed
high predictive power when transferred. The poor transferability of the
Default model is likely explained by default settings overfitting the
model on the training points (Warren et al., 2014). This finding
corroborates other studies showing that high predictive power in the
model training area is not a guarantee of transferability (Duque-Lazo et
al., 2016; Warren et al., 2014). We echo that recommendations of
Shcheglovitova and Anderson (2013) and Muscarella et al. (2014) that
model tuning should be performed to estimate optimal model complexity
when transferring MaxEnt models in space and/or time in order to
maximise their usefulness.
